### 2017-02-02 CS798-ConvexityAndOptimization/assignments/assignment_01.tex
Fixed 3b of apx algs a1
conv opt 2b of a1

>>>>

Let $n \in \N$ and consider the function $f: S^n_{++} \rightarrow \R$ given by
$$f(x) = (\det X)^\frac{1}{n}.$$
We will demonstrate that $f$ is concave. It suffices to show that the function 
$$g(t) = (\det (Z + t V))^\frac{1}{n} $$
is concave for arbitrary $Z \in S^n_{++}$ and $V \in S^n$. Note that $g$ is the restriction of $f$ to some arbitrary line. The domain of $g$ is $\{t \in \R: Z + tV \in S^n_{++}\}$.
Since $Z \in S^n_{++}$ we can decompose $Z$ as $Z=Z^\frac{1}{2}Z^\frac{1}{2}$ with $Z^\frac{1}{2}$ non-singular. Let $\lambda_1, \dots, \lambda_n$ denote the eigenvalues of $Z^\frac{-1}{2}VZ^\frac{-1}{2}$. Then we observe that
$$g(t) = (\det(Z+tV))^\frac{1}{n} = (\det(Z^\frac{1}{2}(I + tZ^\frac{-1}{2}VZ^\frac{-1}{2})Z^\frac{1}{2}))^\frac{1}{n}.$$
Using that $\det(\cdot)$ distributes over matrix products and that $\det(A)$ is equal to the product of the eigenvalues of $A$ we have
$$g(t) = \det(Z^\frac{1}{2})^\frac{2}{n}\det(1 + tZ^\frac{-1}{2}VZ^\frac{-1}{2}) = \det(Z^\frac{1}{2})^\frac{2}{n}(\Pi_{i=1}^n 1+t\lambda_i)^\frac{1}{n}.$$
Now we will show that $g$ is a composition of affine and concave functions, and hence that $g$ is concave. Let $h : \R \rightarrow \R^n$ be the affine function given by
$$h(t)_i = 1+t\lambda_i$$
for $i = 1\dots n$. Let $ y : \R^n \rightarrow \R$ be the geometric mean function
$$y(x) = (\Pi_{i=1}^n x_i)^\frac{1}{n}.$$
We know from the notes that $y$ is concave. Finally, let $z: \R\rightarrow \R$ be the affine (linear actually) function given by
$$z(x) = \det(Z^\frac{1}{2})^\frac{2}{n}$$
Then we see that
$$g(t) = z\circ y\circ h(t).$$
Thus $g$ is a composition of affine functions with a concave function and hence $g$ is concave.$\blacksquare$
